### 2022-02-01 paper/slides/background-ml2.tex
colloquium checkpoint

\begin{frame}
\frametitle{Machine Learning Review: What is Generalization Error?}

Standard Experimental Design:
\begin{itemize}
\item Split data into training/test set
\item Train model on training set
\item Evaluate model on the test set
\end{itemize}

%\includegraphics[width=\textwidth]{img/llustration-of-the-underfitting-overfitting-issue-on-a-simple-regression-case-Data}

\includegraphics[width=\textwidth]{img/over-under-fit}

{
\fontsize{4}{4}
{\tiny Image source:} \url{https://scikit-learn.org/stable/auto_examples/model_selection/plot_underfitting_overfitting.html}
}
\end{frame}