### 2022-02-01 paper/slides/conclusion.tex

\begin{frame}{Summary}

Theory predicts the following generalization errors:
\begin{itemize}
\item cross entropy is $O(\sqrt{k/n})$
\item tree loss is $\begin{cases}O(\sqrt{\log k/n})&c<=1\\O(\sqrt{k^{1-1/c}/n})&c>1\end{cases}$% if $c<=1$ else $O(\sqrt{k^{1-1/c}/n})$
\end{itemize}

\vspace{0.1in}
We hypothesize the optimal is $O(\sqrt{c/n})$

\vspace{0.1in}
Experiments:
\begin{itemize}
\item tree loss outperforms cross entropy loss:
\begin{itemize}
\item in all regimes of ablation study
\item even for bad trees! (lemma proves this is always true)
\end{itemize}
\item good performance on real world problems
\begin{itemize}
\item still working on gettint SOTA on ImageNet...
\end{itemize}
\end{itemize}

\vspace{0.1in}
\textbf{Tree Loss is easy to implement! Code available!}

\url{https://github.com/cora1021/TreeLoss}
\end{frame}