### 2022-02-01 paper/slides/covertree/contributions-merge.tex

\begin{frame}[fragile]{Merging cover trees}

Merging cover trees gives us a parallel tree construction algorithm

\vspace{0.15in}

Sometimes, merging cover trees is \textbf{easy}:

\begin{center}
\begin{tikzpicture}
    [ draw
    , every node/.style={minimum size=10mm,fill=white}
    , level/.style={sibling distance = 23mm/#1, level distance=12mm}
    %,    level distance = 1.5cm}
    , sibling distance=8mm
    ]
\draw (-2.3,0) -- (8.6,0)[dotted];
\draw (-2.3,-12mm) -- (8.6,-12mm)[dotted];
\draw (-2.3,-24mm) -- (8.6,-24mm)[dotted];
\node[shape=circle,draw] at (0,0) {10}
    child { node[circle,draw] {8}
        child { node[circle,draw] {7}  }
        child { node[circle,draw] {9} }
        }
    child [color=white] {}
    %child { node[circle,draw] {12}
        %child { node[circle,draw] {9}  }
        %child { node[circle,draw] {13} }
        %}
    ;
\node[shape=circle,draw,fill=lightgreen] at (5,-12mm) {12}
    %child { node[circle,draw] {8}
        %child { node[circle,draw] {7}  }
        %child { node[circle,draw,fill=lightgreen,line width=1pt] {9} }
        %}
    %child { node[circle,draw] {12}
        child { node[circle,draw] {11}  }
        child { node[circle,draw] {13} }
        %}
    ;
\node[fill=none] at (8,3mm) {level 3};
\node[fill=none] at (8,-9mm) {level 2};
\node[fill=none] at (8,-21mm) {level 1};
\end{tikzpicture}
\end{center}

\vspace{0.1in}
No runtime bound on the merge operation, but it is fast in practice

%\vspace{0.1in}
%But, if the runtime is $o(n)$, then we get an algorithm for tree construction that takes time $O(n)$
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Merging cover trees}

Merging cover trees gives us a parallel tree construction algorithm

\vspace{0.15in}

Sometimes, merging cover trees is \textbf{hard}:

\begin{center}
\begin{tikzpicture}
    [ draw
    , every node/.style={minimum size=10mm,fill=white}
    , level/.style={sibling distance = 23mm/#1, level distance=12mm}
    %,    level distance = 1.5cm}
    , sibling distance=8mm
    ]
\draw (-2.3,0) -- (8.6,0)[dotted];
\draw (-2.3,-12mm) -- (8.6,-12mm)[dotted];
\draw (-2.3,-24mm) -- (8.6,-24mm)[dotted];
\node[shape=circle,draw] at (0,0) {10}
    child { node[circle,draw] {8}
        child { node[circle,draw] {7}  }
        child { node[circle,draw] {9} }
        }
    child [color=white] {}
    %child { node[circle,draw] {12}
        %child { node[circle,draw] {9}  }
        %child { node[circle,draw] {13} }
        %}
    ;
\node[shape=circle,draw,fill=lightred] at (5,-12mm) {11.5}
    %child { node[circle,draw] {8}
        %child { node[circle,draw] {7}  }
        %child { node[circle,draw,fill=lightgreen,line width=1pt] {9} }
        %}
    %child { node[circle,draw] {12}
        child { node[circle,draw] {11}  }
        child { node[circle,draw] {13} }
        %}
    ;
\node[fill=none] at (8,3mm) {level 3};
\node[fill=none] at (8,-9mm) {level 2};
\node[fill=none] at (8,-21mm) {level 1};
\end{tikzpicture}
\end{center}

\vspace{0.1in}
No runtime bound on the merge operation, but it is fast in practice

%\vspace{0.1in}
%But, if the runtime is $o(n)$, then we get an algorithm for tree construction that takes time $O(n)$
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{The effect of parallel tree \emph{construction} on small datasets}
\begin{center}
\graphicspath{{slides/covertree/paperimg/}}
\input{slides/covertree/paperimg/parallel-ancestor-build}
\begin{tikzpicture}
    \node[draw,fill=blue,minimum width=0.05in,minimum height=0.3in] at (0.3in,0) {};
    \node at (0.3in,0.63in) {\small\rotatebox{90}{Our cover tree}};
    \node[minimum width=0.05in,minimum height=0.3in] at (0.45in,0) {};
    \node at (0.45in,0.61in) {\small\rotatebox{90}{}};
    \node at (0,-0.475in) {};
\end{tikzpicture}
\end{center}

\vspace{0.1in}
Experiments run on an Amazon AWS instance with 16 true cores
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Parallel tree construction really matters on larger data sets}

%on small datasets with cheap metrics, parallel construction is less useful

\vspace{0.1in}
On large datasets with an expensive metric, parallelism is more useful

\vspace{0.15in}
Yahoo! Flickr dataset with 1.5 million images and earth mover distance

%\vspace{0.2in}
\vspace{0.15in}
\begin{center}
\Large
\begin{tabular}{ccccc}
\hline
num cores
 & \multicolumn{2}{c}{simplified tree} & \multicolumn{2}{c}{nearest ancestor tree} \\
%& \multicolumn{2}{c}{construction} & \multicolumn{2}{c}{construction} \\ %\cline{2-5}
& time & speedup & time & speedup \\
\hline
\hline
1  & 70.7 min & 1.0 & 210.9 min& 1.0\\
2  & 36.6 min & 1.9 & 94.2 min & 2.2\\
4  & 18.5 min & 3.8 & 48.5 min & 4.3\\
8  & 10.2 min & 6.9 & 25.3 min & 8.3\\
16 & 6.7 min & 10.5 & 12.0 min & 17.6\\
\hline
\end{tabular}
\end{center}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{The effect of parallel tree \emph{construction and query}}
\begin{center}
\graphicspath{{slides/covertree/paperimg/}}
\input{slides/covertree/paperimg/parallel}
\definecolor{colorOrig}{RGB}{102,51,0}
\definecolor{colorMlpack}{RGB}{204,153,0}
\begin{tikzpicture}
    \node[draw,fill=colorOrig,minimum width=0.05in,minimum height=0.3in] at (0,0) {};
    \node at (0,0.79in) {\small\rotatebox{90}{Reference cover tree}};
    \node[draw,fill=colorMlpack,minimum width=0.05in,minimum height=0.3in] at (0.15in,0) {};
    \node at (0.15in,0.80in) {\small\rotatebox{90}{MLPack's cover tree}};
    \node[draw,fill=blue,minimum width=0.05in,minimum height=0.3in] at (0.3in,0) {};
    \node at (0.3in,0.63in) {\small\rotatebox{90}{Our cover tree}};
    \node[minimum width=0.05in,minimum height=0.3in] at (0.45in,0) {};
    \node at (0.45in,0.61in) {\small\rotatebox{90}{}};
    \node at (0,-0.475in) {};
\end{tikzpicture}
\end{center}

\vspace{0.1in}
Experiments run on an Amazon AWS instance with 16 true cores
\end{frame}