### 2022-02-01 paper/slides/covertree/contributions.tex

\begin{frame}[fragile]{My contributions}

\large

I made cover trees faster in practice:
\begin{itemize}

\item Two new cover tree variants:

\begin{itemize}
\item The simplified cover tree

\item The nearest ancestor cover tree
\end{itemize}
\large

\item A cache efficient memory layout for the cover tree

\item A method for parallel tree construction
\end{itemize}

\vspace{0.15in}
I'll demonstrate the improvements empirically
\begin{itemize}
\item Experiments use the \emph{all-nearest-neighbor} query method
\item Datasets:
\begin{itemize}
\item Benchmark data sets using the Euclidean distance
\item The protein data bank using random walk distance
\item Flickr 1.5 million image data set using earth mover's distance
\end{itemize}
\end{itemize}

\end{frame}