### 2022-02-01 paper/slides/covertree/experiment-protein.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]{Experiment: protein data bank data}

%This experiment uses the protein data and the random walk graph kernel.
\vspace{0.1in}
Protein distance measured using the random walk graph kernel
\vspace{0.1in}

{
\begin{center}
\graphicspath{{slides/covertree/paperimg/}}
\input{slides/covertree/paperimg/protein-pres}
\begin{tikzpicture}
    \node[draw,fill=darkred,minimum width=0.03in,minimum height=0.3in] at (0,0) {};
    \node at (0,0.77in) {\small\rotatebox{90}{Original cover tree}};
    \node[draw,fill=darkgreen,minimum width=0.03in,minimum height=0.3in] at (0.15in,0) {};
    \node at (0.15in,0.82in) {\small\rotatebox{90}{Simplified cover tree}};
    %\node[draw,fill=blue,minimum width=0.03in,minimum height=0.3in] at (0.3in,0) {};
    %\node at (0.3in,1.01in) {\small\rotatebox{90}{Nearest ancestor cover tree}};
    %\node[minimum width=0.05in,minimum height=0.5in] at (0.45in,0) {};
    %\node at (0.6in,0.7in) {\small\rotatebox{90}{ dotted lines: construction only}};
    %\node at (0.75in,0.81in) {\small\rotatebox{90}{ solid lines: construction and query}};
    \node at (0,-0.5in) {};
\end{tikzpicture}
\end{center}
}

\end{frame}