### 2022-02-01 paper/slides/covertree/experimental-setup.tex

\begin{frame}{Experimental setup}
\Large

Three data sources:
\begin{itemize}
\item Protein dataset with the random walk graph distance
\item Yahoo! 1.5 million creative common images with the earth movers distance
\item MLPack benchmarks with Euclidean distance
    %\begin{itemize}
    %\item Selected the 8 largest datasets
    %\item Maximum size
    %\end{itemize}
\end{itemize}

\vspace{0.15in}
Benchmarking procedure:
\begin{itemize}
\item Construct a cover tree on the dataset
\item For each data point in the dataset, find the nearest neighbor
\item AWS instance with 16 true cores
\end{itemize}

\end{frame}