### 2022-02-01 paper/slides/covertree/img-allcovertree.tex

\documentclass[tikz]{standalone}
\usetikzlibrary{patterns}
\begin{document}

\definecolor{lightred}{RGB}{255,200,200}
\definecolor{lightgreen}{RGB}{200,255,200}
\definecolor{lightblue}{RGB}{200,200,255}
\definecolor{darkred}{RGB}{127,0,0}
\definecolor{darkgreen}{RGB}{0,127,0}
\definecolor{darkblue}{RGB}{0,0,127}

\definecolor{lightgray}{RGB}{205,210,255}
\definecolor{lightgray2}{RGB}{235,240,255}

\begin{tikzpicture}
    [ draw
    , every node/.style={line width=1pt, minimum size=10mm,fill=lightgray}
    , level 1/.style={sibling distance = 30mm, level distance=20mm}
    , level 2/.style={sibling distance = 24mm, level distance=12mm}
    , level 3/.style={sibling distance = 12mm, level distance=12mm}
    , level 4/.style={sibling distance = 12mm, level distance=20mm}
    , dash/.style={edge from parent/.style={gray,thin,draw}}
    , norm/.style={edge from parent/.style={black,very thick,draw}}
    , funk/.style={preaction={fill=white},pattern=north west lines,pattern color=lightgray,draw=gray,circle,thin}
    %,    level distance = 1.5cm}
    , sibling distance=8mm
    ]
\draw (-4.3,  0mm) -- (18,  0mm)[dotted];
\draw (-4.3,-20mm) -- (18,-20mm)[dotted];
\draw (-4.3,-32mm) -- (18,-32mm)[dotted];
\draw (-4.3,-44mm) -- (18,-44mm)[dotted];
\draw (-4.3,-64mm) -- (18,-64mm)[dotted];
%\node[circle,draw=gray,thin,pattern=north west lines,pattern color=lightgray] at (0,0) {10}
\node[funk] at (0,0) {10}
    child[dash] {
        node[shape=circle,draw] (b1) {10}
        child[norm] { node[circle,draw=black,very thick] {8}
            child[norm] { node[circle,draw] {7}         child[dash] { node[funk] (a) {7} } }
            child[dash] { node[funk] {8}  child[dash] { node[funk] {8} } }
            child[opacity=0] { node[opacity=0] {} }
            }
        child[norm] { node[circle,draw=black] (b2) {10}
            child[norm] { node[circle,draw,fill=lightred] {8.5}         child[dash] { node[funk] {8.5} } }
            child[dash] { node[funk] {10} child[dash] { node[funk] {10}} }
            child[norm] { node[circle,draw,fill=lightred] (c1) {11.5}        child[dash] { node[funk] {11.5}} }
            }
        child[norm] { node[circle,draw=black] {12}
            child[opacity=0] { node[opacity=0] {} }
            child[dash] { node[funk] {12} child[dash] { node[funk] {12}} }
            child[norm] { node[circle,draw] {13}        child[dash] { node[funk] {13}} }
            }
        }
    ;

\node[opacity=0] at (8,0) {}
    child[norm,opacity=0] { node[opacity=1,shape=circle,draw] {10}
        child[opacity=1] { node[circle,draw] {8}
            child { node[circle,draw] {7}  }
            child { node[circle,draw,fill=lightred,line width=1pt] (c2) {11.5} }
            }
        child[opacity=1] { node[circle,draw] {12}
            child { node[circle,draw,fill=lightred,line width=1pt] {8.5}  }
            child { node[circle,draw] {13} }
            }
        }
    ;

\node[opacity=0] at (14.5,0) {}
    child[norm,opacity=0] { node[opacity=1,shape=circle,draw] {10}
        child[opacity=1] { node[circle,draw] {8}
            child { node[circle,draw] {7}  }
            child { node[circle,draw,fill=lightgreen,line width=1pt] {8.5} }
            }
        child[opacity=1] { node[circle,draw] {12}
            child { node[circle,draw,fill=lightgreen,line width=1pt] (d1) {11.5}  }
            child { node[circle,draw] {13} }
            }
        }
    ;

\node[fill=none] at (19,0mm) {level $\infty$};
\node[fill=none] at (19,-10mm) {$\vdots$};
\node[fill=none] at (19,-20mm) {level 2};
\node[fill=none] at (19,-32mm) {level 1};
\node[fill=none] at (19,-44mm) {level 0};
\node[fill=none] at (19,-54mm) {$\vdots$};
\node[fill=none] at (19,-64mm) {level $-\infty$};

\node[fill=none] at (-2,-0.70) {\textbf{original}};
\node[fill=none] at (-2,-1.05) {\textbf{cover tree}};

\node[fill=none] at (8,-0.70) {\textbf{simplified}};
\node[fill=none] at (8,-1.05) {\textbf{cover tree}};

\node[fill=none] at (14.5,-0.70) {\textbf{nearest ancestor}};
\node[fill=none] at (14.5,-1.05) {\textbf{cover tree}};

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\node[fill=none] at (0,-8) {\parbox{2.5in}{
    these ``implicit'' nodes are needed to describe the original cover tree,
    but they don't actually get implemented
    }};
\draw[->,out=180,in=270] (-3.5,-8) to (a);

\node[fill=none] at (3.5,-1.5) {\parbox{2in}{
    once a value appears in the tree, it must be in every lower level
    }};
\draw[->,out=180,in=45] (0.75,-1.5) to (b1);
%\draw[->,out=225,in=45] (0.75,-1.65) to (b2);

\node[fill=none] at (8,-5.75) {\parbox{2.3in}{
    nodes can be placed suboptimally even in one dimension
    }};
\draw[->,out=180,in=315] (4.85,-5.75) to (c1);
\draw[->,out=80,in=270] (7.25,-5.35) to (c2);

\node[fill=none] at (14,-7.25) {\parbox{3in}{
    in one dimension, nearest ancestor trees behave like balanced binary search trees
    }};
\draw[->,out=80,in=270] (14.25,-6.8) to (d1);

\end{tikzpicture}
\end{document}