### 2022-02-01 paper/slides/covertree/img-originalcovertree.tex

\documentclass[tikz]{standalone}
\begin{document}

\definecolor{lightred}{RGB}{255,200,200}
\definecolor{lightgreen}{RGB}{200,255,200}
\definecolor{lightblue}{RGB}{200,200,255}
\definecolor{darkred}{RGB}{127,0,0}
\definecolor{darkgreen}{RGB}{0,127,0}
\definecolor{darkblue}{RGB}{0,0,127}

\definecolor{lightgray}{RGB}{205,210,255}

\begin{tikzpicture}
    [ draw
    , every node/.style={line width=1pt, minimum size=10mm,fill=lightgray}
    , level 1/.style={sibling distance = 30mm, level distance=20mm}
    , level 2/.style={sibling distance = 30mm, level distance=12mm}
    , level 3/.style={sibling distance = 12mm, level distance=12mm}
    , level 4/.style={sibling distance = 12mm, level distance=20mm}
    , dash/.style={edge from parent/.style={gray,thin,draw}}
    , norm/.style={edge from parent/.style={black,very thick,draw}}
    %,    level distance = 1.5cm}
    , sibling distance=8mm
    ]
\draw (-4.3,0mm) -- (5,0mm)[dotted];
\draw (-4.3,-20mm) -- (5,-20mm)[dotted];
\draw (-4.3,-32mm) -- (5,-32mm)[dotted];
\draw (-4.3,-44mm) -- (5,-44mm)[dotted];
\draw (-4.3,-64mm) -- (5,-64mm)[dotted];
\node[circle,draw=gray,thin] at (0,0) {10}
    child[dash] {
        node[shape=circle,draw] {10}
        child[norm] { node[circle,draw=black,very thick] {8}
            child[norm] { node[circle,draw] {7}         child[dash] { node[circle,draw=gray,thin] {7} } }
            child[dash] { node[circle,draw=gray,thin] {8}  child[dash] { node[circle,draw=gray,thin] {8} } }
            }
        child[norm] { node[circle,draw=black] {10}
            child[norm] { node[circle,draw] {9}         child[dash] { node[circle,draw=gray,thin] {9} } }
            child[dash] { node[circle,draw=gray,thin] {10} child[dash] { node[circle,draw=gray,thin] {10}} }
            child[norm] { node[circle,draw] {11}        child[dash] { node[circle,draw=gray,thin] {11}} }
            }
        child[norm] { node[circle,draw=black] {12}
            child[dash] { node[circle,draw=gray,thin] {12} child[dash] { node[circle,draw=gray,thin] {12}} }
            child[norm] { node[circle,draw] {13}        child[dash] { node[circle,draw=gray,thin] {13}} }
            }
        }
    ;
\node[fill=none] at (6,0mm) {level $\infty$};
\node[fill=none] at (6,-10mm) {$\vdots$};
\node[fill=none] at (6,-20mm) {level 2};
\node[fill=none] at (6,-32mm) {level 1};
\node[fill=none] at (6,-44mm) {level 0};
\node[fill=none] at (6,-54mm) {$\vdots$};
\node[fill=none] at (6,-64mm) {level $-\infty$};
\end{tikzpicture}
\end{document}