### 2022-02-01 paper/slides/covertree/img-simplifiedcovertree2.tex

\documentclass[tikz]{standalone}
\begin{document}

\definecolor{lightred}{RGB}{255,200,200}
\definecolor{lightgreen}{RGB}{200,255,200}
\definecolor{lightblue}{RGB}{200,200,255}
\definecolor{darkred}{RGB}{127,0,0}
\definecolor{darkgreen}{RGB}{0,127,0}
\definecolor{darkblue}{RGB}{0,0,127}

\begin{tikzpicture}
    [ draw
    , every node/.style={minimum size=10mm,fill=white}
    , level/.style={sibling distance = 23mm/#1, level distance=12mm}
    %,    level distance = 1.5cm}
    , sibling distance=8mm
    ]
\draw (-2.3,0) -- (8.6,0)[dotted];
\draw (-2.3,-12mm) -- (8.6,-12mm)[dotted];
\draw (-2.3,-24mm) -- (8.6,-24mm)[dotted];
\node[shape=circle,draw] at (0,0) {10}
    child { node[circle,draw] {8}
        child { node[circle,draw] {7}  }
        child { node[circle,draw,fill=lightred,line width=1pt] {11} }
        }
    child { node[circle,draw] {12}
        child { node[circle,draw,fill=lightred,line width=1pt] {9}  }
        child { node[circle,draw] {13} }
        }
    ;
\node[shape=circle,draw] at (5,0) {10}
    child { node[circle,draw] {8}
        child { node[circle,draw] {7}  }
        child { node[circle,draw,fill=lightgreen,line width=1pt] {9} }
        }
    child { node[circle,draw] {12}
        child { node[circle,draw,fill=lightgreen,line width=1pt] {11}  }
        child { node[circle,draw] {13} }
        }
    ;
\node[fill=none] at (8,3mm) {level 3};
\node[fill=none] at (8,-9mm) {level 2};
\node[fill=none] at (8,-21mm) {level 1};
\end{tikzpicture}
\end{document}