### 2022-02-01 paper/slides/results-summary.tex

\begin{frame}{(Informal) Main Result}

Let
\begin{itemize}
%\item $d$ features
\item $k$ be the number of classes, and%\uncover<2>{\textcolor{red}{, doubling dimension $c$}}
\item $n$ be the number of data points.
\end{itemize}

%The generalization error of the standard cross entropy loss decays as
%\begin{equation}
    %O(\sqrt{dk/n}).
%\end{equation}

Then almost all classification problems have
\begin{equation}
    \text{generalization error} = \Theta(\sqrt{k/n})
\end{equation}
in the PAC framework.

\uncover<2>{
    \textcolor{red}{(\emph{new})}
    Let
    \begin{itemize}
        \item $c$ be the doubling dimensionality of the classes.
    \end{itemize}
    Then we probide a simple algorithm with
    \begin{equation}
        \text{generalization error} = 
        \begin{cases}
            O(\sqrt{\log k/n}) & c \le 1 \\
            O(\sqrt{k^{1-c}/n}) & c > 1 \\
        \end{cases}
    \end{equation}
    and we conjecture that $\Theta(\sqrt{c/n})$ is optimal.
        %We provide a simple algorithm with generalization error
    %\begin{equation}
        %O(\sqrt{d\log k}/n)
    %\end{equation}
    }

\end{frame}