### 2022-02-05 paper/slides/conjecture.tex

\begin{frame}
\frametitle{Optimal Generalization Error}

\textbf{Recall:} standard generalization error bound is $\Theta(\sqrt{kd/n})$

\pause
\begin{block}{Conjecture}
If the classes have metric structure with doubling dimension $c$,
then 
\begin{equation}
\text{the optimal generalization error} = \Theta\left(\sqrt{cd/n}\right)
\end{equation}
\end{block}

\pause
Why?
\pause
\begin{itemize}
    \item if the metric is discrete (no assumptions), then $c=\Theta(k)$ recovering known bound
    \pause
\item if the metric is $\mathbb Z$, $k=\infty, c=\Theta(1)$ and the known bound is $O(\sqrt{d/n})$ via Poisson regression
    \pause
\item if the metric is $m$-sphere, then $c=\Theta(m)$ and prior work shows the bound is $O(\sqrt{md/n})$
via MvMF regression \citep{izbicki2019exploiting}
\end{itemize}

\end{frame}