### 2022-02-05 paper/slides/experiments-image.tex

\begin{frame}{Experiments: Real World}

\resizebox{\columnwidth}{!}{
\input{fig/real_world.tex}
}

    \vspace{0.2in}
Experimental Details:

\begin{itemize}
    \item CIFAR100/ImageNet: Image data, 100/1000 classes, ResNet features

        \citep{He2016DeepRL}

    \item Twitter: Text data, 80 classes, BERT features

        \citep{stoikos2020multilingual}

        \begin{itemize}
            \item predict the emoji (detailed sentiment) of a tweet about coronavirus
        \end{itemize}
\end{itemize}

\end{frame}