### 2022-02-05 paper/slides/outline1.tex

\begin{frame}{Summary (so far)}

\begin{itemize}
\item more classes $\Rightarrow$ harder problem
\begin{itemize}
\item $k$: num classes, $d$: num dimensions, $n$: num data points
\end{itemize}
\begin{equation*}
\text{(cross entropy) generalization error}
=
O\left(\sqrt{\frac{kd}{n}}\right)
~~~~~~~~~~
\end{equation*}
\item real world classes are highly structured

\begin{itemize}
\item the doubling dimension $c <\!\!< k$

\end{itemize}
\begin{equation*}
\text{(conjectured optimal) generalization error}
=
O\left(\sqrt{\frac{cd}{n}}\right)
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~.
\end{equation*}

\item \textbf{contribution:} $U$/$V$-tree loss exploits arbitrary metric structure

\begin{equation*}
\text{(tree loss) generalization error} = 
\begin{cases}
\sqrt{\frac{d\log k}{n}} & c\le 1 \\
\sqrt{\frac{dk^{1-1/c}}{n}} & c>1 \\
\end{cases}
\end{equation*}

\item \textbf{bonus:} best performance on synthetic and real world experiments

\end{itemize}

\vspace{-4in}
\begin{tikzpicture}
    \node at (0,0) {};
    \node at (0,4in) {};

    \node[anchor=north west,fill=white,minimum width=4.5in,minimum height=0.7in] at (0,2.94in) {};
    \node[anchor=north west,fill=white,minimum width=4.5in,minimum height=0.74in] at (0,2.04in) {};
    \node[anchor=north west,fill=white,minimum width=4.5in,minimum height=0.7in] at (0,1.04in) {};
    \node[anchor=north west,fill=white,minimum width=4.5in,minimum height=0.3in] at (0,0.3in) {};
\end{tikzpicture}
\end{frame}